Semantic Textual Similarity using deep learning

### 2017-06-03 paper/main.tex

\usepackage[utf8]{inputenc}
\usepackage[pdftex]{graphicx}
\usepackage{booktabs}
\usepackage{amsmath}
\usepackage{amssymb}

@@

\subsection{SVM}

@@

\begin{figure}
\begin{center}
\includegraphics[width=\columnwidth]{only_2nd_layer.png}
\caption{Depiction of blabla}
\label{fig:lstm_2nd_layer}
\end{center}
\end{figure}

@@

\bibliographystyle{tar2016}
\bibliography{tar2016}

>>>>

\usepackage[utf8]{inputenc}
\usepackage[pdftex]{graphicx}
\usepackage{booktabs}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{subcaption}

@@

\subsection{Data analysis}
We used dataset from SemEval 2017 
Task 1\footnote{http://alt.qcri.org/semeval2017/task1/}. 
The dataset has 250 instances. Each instance consists of two sentences and our tash is to predict semantic similarity between them. In order to get some deeper insights on data we did some minor data analysis. We got to conclusion that we are dealing with very short text. Also, we concluded that the smaller difference between lengths of sentences is, the more similar they are. This can be seen clearer on following plots.

\begin{table}
\caption{Stats about lengths of sentences}
\label{tab:narrow-table}
\begin{center}
\begin{tabular}{cll}
\toprule
& Without stopwords & With stopwords \\
\midrule
Sentence length & 33.004 & 43.838 \\
Tokens in sentence & 5.488 & 8.702 \\
\bottomrule
\end{tabular}
\end{center}
\end{table}

\begin{figure}[h]
\begin{center}
\includegraphics[width=\columnwidth]{sen_len_no_stop.png}
\caption{Sentence lengths without stopwords}
\label{fig:lstm_2nd_layer}
\end{center}
\end{figure}

\begin{figure}[h]
\begin{center}
\includegraphics[width=\columnwidth]{sen_len_with_stop.png}
\caption{Sentence lengths with stopwords}
\label{fig:lstm_2nd_layer}
\end{center}
\end{figure}

\begin{figure}[h]
\begin{center}
\includegraphics[width=\columnwidth]{tokens_no_stop.png}
\caption{Number of tokens without stopwords}
\label{fig:lstm_2nd_layer}
\end{center}
\end{figure}

\begin{figure}[h]
\begin{center}
\includegraphics[width=\columnwidth]{tokens_with_stop.png}
\caption{Number of tokens with stopwords}
\label{fig:lstm_2nd_layer}
\end{center}
\end{figure}

\newpage
\subsection{SVM}

\subsubsection{Features}
All of the features we used are from \citep{Saric2012TakeLabSF} paper.
\paragraph{Ngram Overlap \\}
Let S1 and S2 be the sets of consecutive ngrams in the first and the second sentence, respectively. The ngram overlap is computed for unigrams, bigrams, and trigrams. It is defined as follows:
\begin{equation}\label{eq:ngo}
ngo(S_1, S_2) = 2 \cdot \bigg( \frac{|S_1|}{|S_1 \cap S_2|}+\frac{|S_2|}{|S_1 \cap S_2|}\bigg)^{-1}
\end{equation}

The ngram overlap is the harmonic mean of the degree
to which the second sentence covers the first
and vice versa.\citep{Saric2012TakeLabSF}

\paragraph{WordNet-Augmented Word Overlap \\}
In order to determine some semantical meaning from words we define the WordNet augmented coverage $ PWN(\cdot, \cdot) $:
\begin{equation}\label{eq:pwn}
P_{WN}(S_1, S_2) = \frac{1}{|S_2|} \sum_{w \in S_1} score(w, S_2)
\end{equation}

\begin{equation}\label{eq:pwn-score}
score(w, S) = \begin{cases}
1 & \text{if $w \in S $}\\
\max\limits_{w^{\rq} \in S} sim(w, w^{\rq}) & \text{otherwise}
\end{cases}
\end{equation}

where $sim(\cdot, \cdot)$ represents the WordNet path length
similarity. The WordNet-augmented word overlap
feature is defined as a harmonic mean of
$PWN(S_1, S_2)$ and $PWN(S_2, S_1)$.\citep{Saric2012TakeLabSF}

\paragraph{Weighted Word Overlap \\}
We define information contente measure to give some words more importance:
\begin{equation}\label{eq:inf-content}
ic(w) = ln \frac{\sum_{w^{\rq} \in C} freq(w^{\rq}) }{freq(w)}
\end{equation}
where C is the set of words in the corpus and
freq(w) is the frequency of the word w in the corpus. Since we got very small dataset we used \emph{nltk brown} dataset to calculate word frequency distribution.
The weighted word coverage of the second sentence by the first sentence is given by:
\begin{equation}\label{eq:wwc}
wwc(S_1, S_2) = \frac{\sum_{w \in S_1 \cap S_2} ic(w)}{\sum_{w^{\rq} \in S_2} ic(w^{\rq})}
\end{equation}
where $S_1$ and $S_2$ are words in sentences.\\
The \emph{weighted word overlap} between two sentences
is calculated as the harmonic mean of the
$wwc(S_1, S_2)$ and $wwc(S_2, S_1)$.\citep{Saric2012TakeLabSF}

\paragraph{Number of tokens difference \\}
Difference between number of words per sentence is defined as:
\begin{equation}\label{eq:word-diff}
diff(S_1, S_2) = abs(|S_1| - |S_2|)
\end{equation}
where $S_1$ and $S_2$ are words in sentences.
\paragraph{Vector Space Sentence Similarity \\}
We define each sentence as vector $u(\cdot)$ by summing all word embeddings in the sentence S: $ u(S) = \sum_{w \in S} x_w$ where $x_w$ is word embedding for each word. Another similar
representation $u_w(\cdot)$ uses the information content
$ic(w)$ to weigh the word embedding vector of each word
before summation: $u_w(S) = \sum_{w \in S} ic(w) \cdot x_w.$
We use $|cos(u(S_1), u(S_2))|$ and $|cos(u_w(S_1), u_w(S_2))|$ for the vector space sentence similarity features. \citep{Saric2012TakeLabSF}
\paragraph{Shallow NER Features \\}
For this feature we simply count all words that are capitalized.
\paragraph{Numbers Overlap \\}
For numbers overlap we define following three feature: $log(1+|N_1|+|N_2|)$, $2\cdot|N_1 \cap N_2|/(|N_1|+|N_2|)$ and $N_1 \subseteq N_2 \vee N_2 \subseteq N_1$
where $N_1$ and $N_2$ are sets of numbers in two sentences. We treat all numbers as decimal numbers.

\subsubsection{Model}
All features were preprocessed with StandardScaler\footnote{http://scikit-learn.org/stable/modules/generated/sklearn.preprocessing.StandardScaler.html}. 
\subsubsection{Evaluation}

@@

\newpage
\begin{figure}
\begin{center}
\includegraphics[width=\columnwidth]{only_2nd_layer.png}
\caption{Depiction of blabla}
\label{fig:lstm_2nd_layer}
\end{center}
\end{figure}

@@

\bibliographystyle{tar2016}
\bibliography{tar2016}
\nocite{*}
